Fast Regression of the Tritium Breeding Ratio in Tokamak Fusion Reactors

### 2020-06-12 paper/introduction.tex
implement more of Nikos feedback in two passes

We sought to develop an empirical surrogate model for the tritium breeding ratio (TBR) in a Inertial confinement fusion (ICF) nuclear fusion reactor. Our expensive model was a Monte Carlo (MC) neutronics simulation, \textit{Paramak}\footnote{Provided by collaborator Jonathan Shimwell, at UKAEA.}, which returns a prediction of the TBR for a given configuration of a spherical (ICF) reactor.

@@

where T represents tritium and \isotope[7]{Li}, \isotope[6]{Li} are the more and
less common isotopes of lithium, respectively. The TBR is defined as the ratio between tritium generation in the breeding blanket and consumption in the reactor, whose input parameters in the-expensive MC model fall
into two classes. Continuous parameters, such as material thicknesses and
packing fraction, describe the geometry of a given reactor configuration.
\todo{add all the parameters here} Discrete
categorical parameters then specify material selections, such as
coolants, armours, and neutron multipliers. We set out to produce a fast TBR
function which takes these same input parameters and approximates the MC TBR
model with the greatest achievable regression performance.

>>>>

In this work, we develop an empirical surrogate model for the tritium breeding
ratio (TBR) in a Inertial confinement fusion (ICF) nuclear fusion reactor. The
expensive model (which our surrogate model approximates) is a Monte Carlo (MC)
neutronics simulation, \textit{Paramak}\footnote{Provided by collaborator
Jonathan Shimwell, at UKAEA.}\todo{IOP guidelines discourage footnotes}, which returns a prediction of the TBR for a given configuration of a spherical (ICF) reactor.

@@

where T represents tritium and \isotope[7]{Li}, \isotope[6]{Li} are the more and
less common isotopes of lithium, respectively. The TBR is defined as the ratio between tritium generation in the breeding blanket and consumption in the reactor, whose input parameters in the-expensive MC model fall
into two classes. Continuous parameters, such as material thicknesses and
packing fraction, describe the geometry of a given reactor configuration.
\todo{add all the parameters here} Discrete
categorical parameters specify material selections, such as
coolants, armours, and neutron multipliers. We set out to produce a fast TBR
function which takes these same input parameters and approximates the MC TBR
model with the greatest achievable regression performance.

### 2020-06-12 paper/results.tex
implement more of Nikos feedback in two passes

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{exp2_time_vs_reg}
	\caption{\label{fig:exp2-time-vs-reg}Results of experiment~2, plotted analogously
	to~\Fref{fig:exp1-time-vs-reg}.}
\end{figure}

@@

\subsubsection{Model Comparison}
In the fourth experiment, we aim to create models that yield:
(a)~the best regression performance regardless
of other features, (b)~acceptable performance with the shortest mean
prediction time, or (c)~acceptable performance with the smallest training set.
To this end, we trained 8~surrogates that are presented in~\Fref{fig:reg-performance}
and~\Tref{tbl:exp4-detailed-results}.
Having selected ANNs, GBTs, ERTs, RBFs and SVMs based on the results of
experiments~2-3, we utilised the best-performing hyperparameters.
In pursuit of goal~(a), the best approximator (Model~1,
ANN) achieved~$R^2=\num{0.998}$ and mean prediction
time~$\overline{t}_{\text{pred.}}=\SI{1.124}{\micro\second}$. These correspond
to a standard error~$S=\num{0.013}$ and a relative speedup~$\omega=\num{6916416} \times$
with respect to the MC TBR evaluation baseline. Satisfying
goal~(b), the fastest model (Model~2, ANN) achieved~$R^2=\num{0.985}$,
$\overline{t}_{\text{pred.}}=\SI{0.898}{\micro\second}$, $S=\num{0.033}$
and~$\omega=\num{8659251} \times$.
While these surrogates
were trained on the entire available set of~\num{500000} datapoints, to satisfy
goal~(c) we also trained a more simplified model (Model~4, GBT)
that achieved~$R^2=\num{0.913}$,
$\overline{t}_{\text{pred.}}=\SI{6.125}{\micro\second}$, $S=\num{0.072}$ and $\omega=\num{1269777} \times$
with a set of size only~\num{10000}.

\begin{table*}
	\centering
	\sisetup{round-mode=places,round-precision=3,detect-weight=true,detect-family=true}
	\caption{\label{tbl:exp4-detailed-results}Results of experiment~4. Here,
		average values and standard deviations are reported over 5~cross-validation folds,
		$|\mathcal{T}|$~denotes cross-validation set size ($\times 10^3$)
		and $\omega$ is a relative speedup with respect to
		$\overline{t}_{\text{eval.}}=\SI{7.777049573054314}{\second}$
		measured in the MC TBR model during run~1 (see~\Tref{tbl:sampling-runs}).
	The best-performing method(s) under each are highlighted in bold.}
	\setlength\tabcolsep{4pt}
	\begin{indented}
	\item[]
		\scriptsize
		\begin{tabular}{lrrrrrrrr}
		\toprule
		{} & {} & \multicolumn{4}{c}{Regression performance} &
		\multicolumn{3}{c}{Complexity}\\
		\cmidrule(lr){3-6}
		\cmidrule(lr){7-9}
		Model & $|\mathcal{T}|$ & MAE [TBR] & $S$ [TBR] & $R^2$ [rel.] & $R^2_{\text{adj.}}$ [rel.]
						& $\overline{t}_{\text{trn.}}$ [\si{\milli\second}] &
		$\overline{t}_{\text{pred.}}$ [\si{\milli\second}] & $\omega$ [rel.]\\
		\midrule
		\input{tbl/table11}
		\bottomrule
		\end{tabular}
	\end{indented}
\end{table*}

@@

	\caption{Regression performance of Models 1-8 (from left to right, top to
		bottom) trained in the model comparison experiment, viewed
		as true vs.~predicted TBR on a test set of a selected cross-validation
		fold. Points are coloured by density.}
	\label{fig:reg-performance}
\end{figure}

>>>>

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{exp2_time_vs_reg}
	\caption{\label{fig:exp2-time-vs-reg}Results of Experiment~2, plotted analogously
	to~\Fref{fig:exp1-time-vs-reg}.}
\end{figure}

@@

\subsubsection{Model Comparison}
In Experiment~4, we aim to create models that yield:
(a)~the best regression performance regardless
of other features, (b)~acceptable performance with the shortest mean
prediction time, or (c)~acceptable performance with the smallest training set.
To this end, we trained 8~surrogates that are presented in~\Fref{fig:reg-performance}
and~\Tref{tbl:exp4-detailed-results}.
Having selected ANNs, GBTs, ERTs, RBFs and SVMs based on the results of
Experiments~2 \&~3, we utilised the best-performing hyperparameters.
In pursuit of goal~(a), the best approximator (Model~1,
ANN) achieved~$R^2=\num{0.998}$ and mean prediction
time~$\overline{t}_{\text{pred.}}=\SI{1.124}{\micro\second}$. These correspond
to a standard error~$S=\num{0.013}$ and a relative speedup~$\omega=\num{6916416} \times$
with respect to the MC TBR evaluation baseline. Satisfying
goal~(b), the fastest model (Model~2, ANN) achieved~$R^2=\num{0.985}$,
$\overline{t}_{\text{pred.}}=\SI{0.898}{\micro\second}$, $S=\num{0.033}$
and~$\omega=\num{8659251} \times$.
While these surrogates
were trained on the entire available set of~\num{500000} datapoints, to satisfy
goal~(c) we also trained a more simplified model (Model~4, GBT)
that achieved~$R^2=\num{0.913}$,
$\overline{t}_{\text{pred.}}=\SI{6.125}{\micro\second}$, $S=\num{0.072}$ and $\omega=\num{1269777} \times$
with a set of size only~\num{10000}.

\begin{table*}
	\centering
	\sisetup{round-mode=places,round-precision=3,detect-weight=true,detect-family=true}
	\caption{\label{tbl:exp4-detailed-results}Results of Experiment~4. Here,
		average values and standard deviations are reported over 5~cross-validation folds,
		$|\mathcal{T}|$~denotes cross-validation set size ($\times 10^3$)
		and $\omega$ is a relative speedup with respect to
		$\overline{t}_{\text{eval.}}=\num{7.777049573054314} \pm
		\num{2.8103592103930337} \text{ s}$
		measured in the MC TBR model over~\num{500000} samples.
	The best-performing method(s) under each are highlighted in bold.}
	\setlength\tabcolsep{4pt}
	\begin{indented}
	\item[]
		\scriptsize
		\begin{tabular}{lrrrrrrrr}
		\toprule
		{} & {} & \multicolumn{4}{c}{Regression performance} &
		\multicolumn{3}{c}{Complexity}\\
		\cmidrule(lr){3-6}
		\cmidrule(lr){7-9}
		Model & $|\mathcal{T}|$ & MAE [TBR] & $S$ [TBR] & $R^2$ [rel.] & $R^2_{\text{adj.}}$ [rel.]
						& $\overline{t}_{\text{trn.}}$ [\si{\milli\second}] &
		$\overline{t}_{\text{pred.}}$ [\si{\milli\second}] & $\omega$ [rel.]\\
		\midrule
		\input{tbl/table11}
		\bottomrule
		\end{tabular}
	\end{indented}
\end{table*}

@@

	\caption{Regression performance of Models~1-8 (from left to right, top to
		bottom) trained in Experiment~4, viewed
		as true vs.~predicted TBR on a test set of a selected cross-validation
		fold. Points are coloured by density.}
	\label{fig:reg-performance}
\end{figure}
